---
title: "JPE Dashboard — User Research Feedback Report"
subtitle: "Compiled from [Google Forms Survey Responses](https://docs.google.com/forms/d/e/1FAIpQLSe3M5DvdxeucI8y7Zg3buI1ZdN5KMstl5CULJObjRpkzUfg8Q/viewform?usp=header)"
date: today
date-format: "MMMM YYYY"

format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 3
    toc-title: "Contents"
    toc-location: left
    number-sections: true
    fig-width: 8
    fig-height: 5
    fig-dpi: 150
    embed-resources: true        # single self-contained HTML file
    code-fold: true              # hide code by default, togglable
    code-summary: "Show code"

  pdf:
    toc: true
    number-sections: true
    colorlinks: true
    fig-pos: "H"

execute:
  echo: false
  warning: false
  message: false
  cache: false   # set true to avoid re-fetching sheet on every render
---

```{r setup}
#| label: setup
#| include: false

# ── Install / load packages ──────────────────────────────────────────────────
library("googlesheets4")
library("tidyverse")
library("scales")
library("knitr")
library("kableExtra")
library("fmsb")
library("glue")
library("janitor")
library("lubridate")

# ── Design tokens ────────────────────────────────────────────────────────────
OCEAN <- "#2563EB"; TEAL <- "#0EA5E9"; JADE <- "#10B981"
AMBER <- "#F59E0B"; ROSE <- "#F43F5E"; SLATE <- "#64748B"; INK <- "#1A1F2E"

# ── ggplot theme ─────────────────────────────────────────────────────────────
theme_survey <- function(base_size = 12) {
  theme_minimal(base_size = base_size) %+replace% theme(
    plot.title       = element_text(face = "bold", size = base_size + 2,
                                    color = INK, hjust = 0),
    plot.subtitle    = element_text(size = base_size - 1, color = SLATE,
                                    hjust = 0, margin = margin(b = 8)),
    axis.text        = element_text(color = SLATE, size = base_size - 2),
    axis.title       = element_text(color = SLATE, size = base_size - 1),
    panel.grid.major = element_line(color = "#E2E8F0"),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background  = element_rect(fill = "white", color = NA),
    legend.position  = "bottom",
    legend.title     = element_text(face = "bold", size = base_size - 1),
    strip.text       = element_text(face = "bold", color = INK)
  )
}
theme_set(theme_survey())
```

```{r load-data}
#| label: load-data
#| include: false

# ╔══════════════════════════════════════════════════════════════════════════╗
# ║  CONFIGURE YOUR GOOGLE SHEETS CONNECTION                                ║
# ╠══════════════════════════════════════════════════════════════════════════╣
# ║  Step 1 — Paste your Sheet URL or Sheet ID below.                       ║
# ║    • The URL looks like:                                                 ║
# ║      https://docs.google.com/spreadsheets/d/SHEET_ID/edit               ║
# ║    • Or just use the SHEET_ID string directly.                          ║
# ║                                                                          ║
# ║  Step 2 — Choose AUTH_MODE:                                              ║
# ║    "public"      Sheet is shared "Anyone with link can view"            ║
# ║                  → no login needed, works in CI/CD pipelines             ║
# ║    "interactive" Opens a browser popup to sign in once;                 ║
# ║                  token is cached at ~/.cache/gargle/                     ║
# ║    "service"     Use a service-account JSON key (headless/server use)   ║
# ║                  → set SERVICE_ACCOUNT_JSON path below                  ║
# ╚══════════════════════════════════════════════════════════════════════════╝

SHEET_URL            <- "https://docs.google.com/spreadsheets/d/1wKb_0v9vsbBZdtfBAZeUZTlZqvGR9YnBmRjPCE5Knq8/edit?gid=1013775912#gid=1013775912"
AUTH_MODE            <- "public"      # "public" | "interactive" | "service"
SERVICE_ACCOUNT_JSON <- NULL          # e.g. "credentials/sa_key.json"
SHEET_TAB            <- 1             # 1 = first tab, or use tab name as string

# ── Authenticate ─────────────────────────────────────────────────────────────
if (AUTH_MODE == "public") {
  gs4_deauth()
} else if (AUTH_MODE == "interactive") {
  gs4_auth()
} else if (AUTH_MODE == "service") {
  gs4_auth(path = SERVICE_ACCOUNT_JSON)
}

# ── Read sheet ────────────────────────────────────────────────────────────────
raw <- read_sheet(SHEET_URL, sheet = SHEET_TAB)

# ── Clean column names ────────────────────────────────────────────────────────
df <- raw |>
  clean_names() |>
  mutate(across(where(is.character), str_squish))

# ── Fuzzy-match key columns → standard names ──────────────────────────────────
# Adjust the regex patterns here if your form question wording differs.
safe_rename <- function(df, new_name, pattern) {
  matched <- names(df)[str_detect(names(df), regex(pattern, ignore_case = TRUE))]
  if (length(matched) > 0 && !new_name %in% names(df))
    df <- df |> rename(!!new_name := all_of(matched[[1]]))
  df
}

df <- df |>
  safe_rename("role",           "role|position|title") |>
  safe_rename("wq_familiar",    "water.quality.*familiar|familiar.*water") |>
  safe_rename("gen_familiar",   "genetic.*familiar|familiar.*genetic") |>
  safe_rename("used_dashboard", "used.*dashboard|dashboard.*before") |>
  safe_rename("nav_ease",       "easy.*navig|navig.*easy|first.use") |>
  safe_rename("intuit_buttons", "buttons.*controls|controls.*buttons") |>
  safe_rename("intuit_viz",     "visualization") |>
  safe_rename("intuit_layout",  "layout") |>
  safe_rename("confusing",      "confus|difficult.*use") |>
  safe_rename("improvements",   "improve|features") |>
  safe_rename("add_comments",   "additional.*comment|^comment") |>
  safe_rename("task_chla",      "chlorophyll|chla") |>
  safe_rename("task_do_temp",   "dissolved.*oxygen|do.*temp") |>
  safe_rename("task_spring",    "spring.*run|chinook") |>
  safe_rename("task_creek",     "battle.*creek|creek.*battle|genotype")

# ── Coerce ratings to numeric ─────────────────────────────────────────────────
rating_cols <- c("wq_familiar", "gen_familiar", "nav_ease",
                 "intuit_buttons", "intuit_viz", "intuit_layout")

df <- df |>
  mutate(across(any_of(rating_cols), as.numeric))

n_resp <- nrow(df)
```

```{r helpers}
#| include: false

# Classify task response text as Easy / Difficult / Very Difficult
parse_task <- function(x) {
  x_low <- str_to_lower(x)
  case_when(
    str_detect(x_low, "very.difficult|very hard|couldn|not able") ~ "Very Difficult",
    str_detect(x_low, "difficult|hard")                           ~ "Difficult",
    str_detect(x_low, "easy|straightforward|simple|no problem")   ~ "Easy",
    is.na(x) | str_length(x) < 2                                  ~ NA_character_,
    .default = "Not answered"
  )
}
```

---

## Executive Summary

```{r exec-summary}
#| results: asis

avg_b <- if ("intuit_buttons" %in% names(df)) round(mean(df$intuit_buttons, na.rm = TRUE), 1) else "N/A"
avg_v <- if ("intuit_viz"     %in% names(df)) round(mean(df$intuit_viz,     na.rm = TRUE), 1) else "N/A"
avg_l <- if ("intuit_layout"  %in% names(df)) round(mean(df$intuit_layout,  na.rm = TRUE), 1) else "N/A"

cat(glue(
  "**{n_resp} subject-matter experts** participated in a usability study of the JPE Dashboard, ",
  "which provides interactive access to discrete water quality and fish genetics datasets. ",
  "Respondents represented a range of roles including environmental scientists, ",
  "a JPE data manager, a genetics specialist, and university faculty.\n\n",
  "The dashboard received strong marks for intuitiveness — buttons & controls ({avg_b}/5), ",
  "visualizations ({avg_v}/5), and layout ({avg_l}/5). Navigation on first use was rated ",
  "easy or very easy by all respondents. Task completion was largely successful. ",
  "Key improvement areas centre on controls refinement, data context labelling, ",
  "genetics module expansion, and condensing the Overview page."
))
```

### Prioritized Recommendations

```{r recommendations}
recs <- tribble(
  ~Priority, ~Recommendation,
  "High",   "Recommendation #1",
  "High",   "Recommendation #2",
  "Medium", "Recommendation #3",
  "Medium", "Recommendation #4",
  "Low",    "Recommendation #5",
  "Low",    "Recommendation #6",
  "Low",    "Recommendation #7"
)

recs |>
  mutate(Priority = cell_spec(Priority, color = "white",
    background = case_when(
      Priority == "High"   ~ ROSE,
      Priority == "Medium" ~ AMBER,
      Priority == "Low"    ~ JADE), bold = TRUE)) |>
  kbl(escape = FALSE, col.names = c("Priority", "Recommendation"), align = "ll") |>
  kable_styling(bootstrap_options = c("hover", "condensed"), full_width = TRUE) |>
  column_spec(1, width = "7em", extra_css = "text-align:center") |>
  column_spec(2, width = "60em")
```

---

## Respondent Overview

### Roles

```{r role-chart}
#| fig-cap: "Number of respondents by self-reported role"
if ("role" %in% names(df)) {
  df |>
    count(role) |>
    mutate(role = fct_reorder(str_wrap(role, 35), n)) |>
    ggplot(aes(n, role)) +
    geom_col(fill = OCEAN, width = 0.6) +
    geom_text(aes(label = n), hjust = -0.3, size = 3.5, color = INK, fontface = "bold") +
    scale_x_continuous(expand = expansion(mult = c(0, 0.18)),
                       breaks = scales::breaks_width(1)) +
    labs(title = "Respondent Roles", x = "Count", y = NULL)
}
```


### Dataset Familiarity

```{r familiarity-chart}
#| fig-cap: "Self-reported familiarity (1 = not familiar, 3 = very familiar)"
fam_avail <- intersect(c("wq_familiar", "gen_familiar"), names(df))

if (length(fam_avail) > 0) {
  df |>
    mutate(respondent = row_number()) |>
    select(respondent, any_of(c("wq_familiar", "gen_familiar"))) |>
    pivot_longer(-respondent, names_to = "dataset", values_to = "score") |>
    mutate(
      dataset    = recode(dataset, wq_familiar = "Water Quality",
                                   gen_familiar = "Fish Genetics"),
      respondent = factor(paste0("R", respondent))
    ) |>
    ggplot(aes(respondent, score, fill = dataset)) +
    geom_col(position = position_dodge(0.7), width = 0.6) +
    scale_fill_manual(values = c("Water Quality" = OCEAN, "Fish Genetics" = JADE),
                      name = NULL) +
    scale_y_continuous(breaks = 1:3,
                       labels = c("1 – Not familiar", "2 – Somewhat", "3 – Very familiar"),
                       limits = c(0, 3.5)) +
    labs(title = "Dataset Familiarity by Respondent", x = NULL, y = NULL)
}
```

### Prior Dashboard Experience

```{r dashboard-exp}
if ("used_dashboard" %in% names(df)) {
  df |>
    count(used_dashboard) |>
    mutate(Percent = scales::percent(n / sum(n), accuracy = 1)) |>
    kbl(col.names = c("Used dashboards before?", "Count", "Percent"), align = "lrr") |>
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  full_width = FALSE)
}
```

---

## Usability Ratings

### First-Use Navigation Ease

```{r nav-ease}
#| fig-cap: "Distribution of first-use navigation ease scores"
if ("nav_ease" %in% names(df)) {
  df |>
    filter(!is.na(nav_ease)) |>
    count(nav_ease) |>
    mutate(label = case_when(
      nav_ease == 1 ~ "1 – Very Easy",
      nav_ease == 2 ~ "2 – Easy",
      nav_ease == 3 ~ "3 – Moderate",
      nav_ease == 4 ~ "4 – Difficult",
      nav_ease == 5 ~ "5 – Very Difficult",
      .default = as.character(nav_ease)
    )) |>
    ggplot(aes(fct_reorder(label, nav_ease), n)) +
    geom_col(fill = TEAL, width = 0.5) +
    geom_text(aes(label = n), vjust = -0.4, size = 4.5,
              fontface = "bold", color = INK) +
    scale_y_continuous(expand = expansion(mult = c(0, 0.2)),
                       breaks = scales::breaks_width(1)) +
    labs(title = "Ease of First-Time Navigation",
         subtitle = "All respondents rated navigation easy or very easy",
         x = NULL, y = "Respondents")
}
```

### Intuitiveness

```{r intuit-bar}
#| fig-cap: "Average intuitiveness scores (1–5 scale)"
intuit_avail <- intersect(c("intuit_buttons", "intuit_viz", "intuit_layout"), names(df))

if (length(intuit_avail) > 0) {
  df |>
    summarise(across(any_of(intuit_avail), ~ mean(.x, na.rm = TRUE))) |>
    pivot_longer(everything(), names_to = "dim", values_to = "avg") |>
    mutate(
      dim   = recode(dim,
                intuit_buttons = "Buttons & Controls",
                intuit_viz     = "Visualizations",
                intuit_layout  = "Layout"),
      color = case_when(avg >= 4 ~ JADE, avg >= 3 ~ AMBER, .default = ROSE)
    ) |>
    ggplot(aes(fct_reorder(dim, avg), avg, fill = color)) +
    geom_col(width = 0.5, show.legend = FALSE) +
    geom_text(aes(label = round(avg, 1)), hjust = -0.3,
              size = 5, fontface = "bold", color = INK) +
    scale_fill_identity() +
    scale_y_continuous(limits = c(0, 5.8), breaks = 0:5) +
    coord_flip() +
    labs(title = "Average Intuitiveness Ratings",
         subtitle = "Green ≥ 4  |  Amber ≥ 3  |  Red < 3",
         x = NULL, y = "Average Score (1–5)")
}
```

### Intuitiveness

```{r intuit-radar}
#| fig-cap: "Radar chart of intuitiveness dimensions"
if (length(intuit_avail) >= 3) {
  avgs <- df |>
    summarise(across(any_of(intuit_avail), ~ mean(.x, na.rm = TRUE))) |>
    rename_with(~ recode(.x,
      intuit_buttons = "Buttons &\nControls",
      intuit_viz     = "Visualizations",
      intuit_layout  = "Layout"
    ))

  radar_df <- rbind(rep(5, ncol(avgs)), rep(1, ncol(avgs)), avgs)

  par(mar = c(1, 1, 2, 1))
  radarchart(radar_df,
    axistype = 1,
    pcol = OCEAN,  pfcol = scales::alpha(OCEAN, 0.25), plwd = 2,
    cglcol = "#E2E8F0", cglty = 1, cglwd = 0.8,
    axislabcol = SLATE, vlcex = 0.9,
    title = "Intuitiveness — Radar Overview"
  )
}
```

### Ratings Summary Table

```{r rating-table}
if (length(intuit_avail) > 0) {
  df |>
    select(any_of(intuit_avail)) |>
    summarise(across(everything(), list(
      Mean = ~ round(mean(.x, na.rm = TRUE), 2),
      SD   = ~ round(sd(.x,   na.rm = TRUE), 2),
      Min  = ~ min(.x, na.rm = TRUE),
      Max  = ~ max(.x, na.rm = TRUE)
    ))) |>
    pivot_longer(everything(),
                 names_to  = c("Dimension", ".value"),
                 names_sep = "_") |>
    mutate(Dimension = recode(Dimension,
      intuit.buttons = "Buttons & Controls",
      intuit.viz     = "Visualizations",
      intuit.layout  = "Layout"
    )) |>
    kbl(align = "lrrrr") |>
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  full_width = FALSE) |>
    column_spec(2, bold = TRUE, color = OCEAN)
}
```

---

## Task Completion

```{r task-chart}
#| fig-cap: "Self-reported ease of completing four guided analytical tasks"
task_avail <- intersect(
  c("task_chla", "task_do_temp", "task_spring", "task_creek"), names(df))

task_labels <- c(
  task_chla    = "Chlorophyll-a\nPeaks (WQ)",
  task_do_temp = "DO vs Temp\nCorrelation (WQ)",
  task_spring  = "Spring Run\nProportion (Genetics)",
  task_creek   = "Battle vs Clear\nCreek (Genetics)"
)

if (length(task_avail) > 0) {
  df |>
    select(any_of(task_avail)) |>
    pivot_longer(everything(), names_to = "task", values_to = "response") |>
    filter(!is.na(response), str_length(response) > 1) |>
    mutate(
      difficulty = parse_task(response),
      task_label = recode(task, !!!task_labels[task_avail]),
      difficulty = factor(difficulty,
        levels = c("Easy", "Difficult", "Very Difficult", "Not answered"))
    ) |>
    filter(!is.na(difficulty)) |>
    count(task_label, difficulty) |>
    ggplot(aes(task_label, n, fill = difficulty)) +
    geom_col(width = 0.55) +
    scale_fill_manual(
      values = c(Easy = JADE, Difficult = AMBER,
                 `Very Difficult` = ROSE, `Not answered` = "#CBD5E1"),
      name = "Difficulty"
    ) +
    scale_y_continuous(breaks = scales::breaks_width(1),
                       expand = expansion(mult = c(0, 0.12))) +
    labs(title = "Task Completion Difficulty",
         subtitle = "Based on respondents' self-reported ease",
         x = NULL, y = "Respondents")
}
```

---

## Qualitative Feedback

### Confusing or Difficult Elements

```{r confusing-table}
if ("confusing" %in% names(df)) {
  df |>
    select(any_of("role"), confusing) |>
    filter(!is.na(confusing), str_length(confusing) > 3) |>
    mutate(confusing = str_trunc(confusing, 450)) |>
    kbl(col.names = c("Role", "Response"), align = "ll") |>
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE) |>
    column_spec(1, width = "18em", italic = TRUE, color = SLATE) |>
    column_spec(2, width = "60em")
}
```

### Feature Requests & Improvements

```{r improvements-table}
if ("improvements" %in% names(df)) {
  df |>
    select(any_of("role"), improvements) |>
    filter(!is.na(improvements), str_length(improvements) > 3) |>
    mutate(improvements = str_trunc(improvements, 450)) |>
    kbl(col.names = c("Role", "Response"), align = "ll") |>
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE) |>
    column_spec(1, width = "18em", italic = TRUE, color = SLATE) |>
    column_spec(2, width = "60em")
}
```

### Additional Comments

```{r comments-table}
if ("add_comments" %in% names(df)) {
  df |>
    select(any_of("role"), add_comments) |>
    filter(!is.na(add_comments), str_length(add_comments) > 3) |>
    mutate(add_comments = str_trunc(add_comments, 450)) |>
    kbl(col.names = c("Role", "Response"), align = "ll") |>
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE) |>
    column_spec(1, width = "18em", italic = TRUE, color = SLATE) |>
    column_spec(2, width = "60em")
}
```



```{r themes-table}
### Recurring Themes
# themes_df <- tribble(
#   ~Theme,                           ~Frequency, ~Sentiment, ~Summary,
#   "Navigation & Controls Friction",  "4 / 6",  "Mixed",    "Tab placement, year slider, and disappearing dropdowns caused friction.",
#   "Data Context & Transparency",     "3 / 6",  "Neutral",  "Requests for aggregation labels, persistent tooltips, and inactive-site messages.",
#   "Genetics Module Expansion",       "3 / 6",  "Positive", "Desire for flow overlays, daily-level counts, raw fish numbers, and upstream stations.",
#   "Verbose Overview Page",           "2 / 6",  "Neutral",  "Overview text is too dense; users prefer concise headings and bullet structure.",
#   "Strong Overall Satisfaction",     "4 / 6",  "Positive", "Multiple respondents praised multi-site comparison and overall ease of use."
# )
# 
# themes_df |>
#   mutate(Sentiment = cell_spec(Sentiment, color = "white",
#     background = case_when(
#       Sentiment == "Positive" ~ JADE,
#       Sentiment == "Negative" ~ ROSE,
#       .default = SLATE
#     ), bold = TRUE)) |>
#   kbl(escape = FALSE, align = "llll") |>
#   kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE) |>
#   column_spec(1, bold = TRUE, width = "22em") |>
#   column_spec(2, width = "8em", extra_css = "text-align:center") |>
#   column_spec(3, width = "9em", extra_css = "text-align:center")
```

---


## Appendix

### User Survey

[Link to survey]

```{r raw-data}
# df |>
#   select(-matches("email|timestamp", ignore.case = TRUE)) |>
#   head(10) |>
#   kbl() |>
#   kable_styling(
#     bootstrap_options = c("striped", "hover", "condensed", "responsive"),
#     full_width = TRUE, font_size = 11
#   ) |>
#   scroll_box(width = "100%", height = "360px")
```

