---
title: "Downstream Dashboard Dashboard — User Research Feedback Report"
subtitle: "Compiled from [Google Forms Survey Responses](https://docs.google.com/forms/d/e/1FAIpQLSe3M5DvdxeucI8y7Zg3buI1ZdN5KMstl5CULJObjRpkzUfg8Q/viewform?usp=header). \n\n **DRAFT - in development**"
date: today
date-format: "MMMM YYYY"

format:
  html:
    theme: cosmo
    checkbox: true
    toc: true
    toc-depth: 3
    toc-title: "Contents"
    toc-location: left
    number-sections: true
    fig-width: 8
    fig-height: 5
    fig-dpi: 150
    embed-resources: true        # single self-contained HTML file
    code-fold: true              # hide code by default, togglable
    code-summary: "Show code"

  pdf:
    toc: true
    number-sections: true
    colorlinks: true
    fig-pos: "H"

execute:
  echo: false
  warning: false
  message: false
  cache: false   # set true to avoid re-fetching sheet on every render
---

```{r setup}
#| label: setup
#| include: false

# ── Install / load packages ──────────────────────────────────────────────────
library("googlesheets4")
library("tidyverse")
library("scales")
library("knitr")
library("kableExtra")
library("fmsb")
library("glue")
library("janitor")
library("lubridate")
library(readxl)
library(here)
# ── Design tokens ────────────────────────────────────────────────────────────
OCEAN <- "#2563EB"; TEAL <- "#0EA5E9"; JADE <- "#10B981"
AMBER <- "#F59E0B"; ROSE <- "#F43F5E"; SLATE <- "#64748B"; INK <- "#1A1F2E"

# ── ggplot theme ─────────────────────────────────────────────────────────────
theme_survey <- function(base_size = 12) {
  theme_minimal(base_size = base_size) %+replace% theme(
    plot.title       = element_text(face = "bold", size = base_size + 2,
                                    color = INK, hjust = 0),
    plot.subtitle    = element_text(size = base_size - 1, color = SLATE,
                                    hjust = 0, margin = margin(b = 8)),
    axis.text        = element_text(color = SLATE, size = base_size - 2),
    axis.title       = element_text(color = SLATE, size = base_size - 1),
    panel.grid.major = element_line(color = "#E2E8F0"),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background  = element_rect(fill = "white", color = NA),
    legend.position  = "bottom",
    legend.title     = element_text(face = "bold", size = base_size - 1),
    strip.text       = element_text(face = "bold", color = INK)
  )
}
theme_set(theme_survey())
```

```{r load-data}
#| label: load-data
#| include: false


SHEET_URL            <- "https://docs.google.com/spreadsheets/d/1wKb_0v9vsbBZdtfBAZeUZTlZqvGR9YnBmRjPCE5Knq8/edit?gid=1013775912#gid=1013775912"
AUTH_MODE            <- "public"      # "public" | "interactive" | "service"
SERVICE_ACCOUNT_JSON <- NULL          
SHEET_TAB            <- 1           

# ── Authenticate ─────────────────────────────────────────────────────────────
if (AUTH_MODE == "public") {
  gs4_deauth()
} else if (AUTH_MODE == "interactive") {
  gs4_auth()
} else if (AUTH_MODE == "service") {
  gs4_auth(path = SERVICE_ACCOUNT_JSON)
}

# ── Read sheet ────────────────────────────────────────────────────────────────
raw <- read_sheet(SHEET_URL, sheet = SHEET_TAB)

# ── Clean column names ────────────────────────────────────────────────────────
df <- raw |>
  clean_names() |>
  mutate(across(where(is.character), str_squish))

# ── Fuzzy-match key columns → standard names ──────────────────────────────────
# Adjust the regex patterns here if your form question wording differs.
safe_rename <- function(df, new_name, pattern) {
  matched <- names(df)[str_detect(names(df), regex(pattern, ignore_case = TRUE))]
  if (length(matched) > 0 && !new_name %in% names(df))
    df <- df |> rename(!!new_name := all_of(matched[[1]]))
  df
}

df <- df |>
  safe_rename("role",           "role|position|title") |>
  safe_rename("wq_familiar",    "water.quality.*familiar|familiar.*water") |>
  safe_rename("gen_familiar",   "genetic.*familiar|familiar.*genetic") |>
  safe_rename("used_dashboard", "used.*dashboard|dashboard.*before") |>
  safe_rename("nav_ease",       "easy.*navig|navig.*easy|first.use") |>
  safe_rename("intuit_buttons", "buttons.*controls|controls.*buttons") |>
  safe_rename("intuit_viz",     "visualization") |>
  safe_rename("intuit_layout",  "layout") |>
  safe_rename("confusing",      "confus|difficult.*use") |>
  safe_rename("improvements",   "improve|features") |>
  safe_rename("add_comments",   "additional.*comment|^comment") |>
  safe_rename("task_chla",      "chlorophyll|chla") |>
  safe_rename("task_do_temp",   "dissolved.*oxygen|do.*temp") |>
  safe_rename("task_spring",    "spring.*run|chinook") |>
  safe_rename("task_creek",     "battle.*creek|creek.*battle|genotype")

# ── Coerce ratings to numeric ─────────────────────────────────────────────────
rating_cols <- c("wq_familiar", "gen_familiar", "nav_ease",
                 "intuit_buttons", "intuit_viz", "intuit_layout")

df <- df |>
  mutate(across(any_of(rating_cols), as.numeric))

n_resp <- nrow(df)
```

```{r helpers}
#| include: false

# Classify task response text as Easy / Difficult / Very Difficult
parse_task <- function(x) {
  x_low <- str_to_lower(x)
  case_when(
    str_detect(x_low, "very.difficult|very hard|couldn|not able") ~ "Very Difficult",
    str_detect(x_low, "difficult|hard")                           ~ "Difficult",
    str_detect(x_low, "easy|straightforward|simple|no problem")   ~ "Easy",
    is.na(x) | str_length(x) < 2                                  ~ NA_character_,
    .default = "Not answered"
  )
}
```

---

## Executive Summary

```{r exec-summary}
#| results: asis

avg_b <- if ("intuit_buttons" %in% names(df)) round(mean(df$intuit_buttons, na.rm = TRUE), 1) else "N/A"
avg_v <- if ("intuit_viz"     %in% names(df)) round(mean(df$intuit_viz,     na.rm = TRUE), 1) else "N/A"
avg_l <- if ("intuit_layout"  %in% names(df)) round(mean(df$intuit_layout,  na.rm = TRUE), 1) else "N/A"
n_resp <- nrow(df)

# Precompute nav ease summary
nav_summary <- if ("nav_ease" %in% names(df)) {
  n_easy <- sum(df$nav_ease <= 2, na.rm = TRUE)
  glue("{n_easy} of {n_resp} participants rated it easy or very easy")
} else "straightforward"

# Precompute task completion summary
task_cols_avail <- intersect(c("task_chla", "task_do_temp", "task_spring", "task_creek"), names(df))

n_easy_tasks <- if (length(task_cols_avail) > 0) {
  df |>
    select(all_of(task_cols_avail)) |>
    pivot_longer(everything(), values_to = "response") |>
    filter(!is.na(response), str_length(response) > 1) |>
    mutate(easy = str_detect(str_to_lower(response), "easy|straightforward|simple")) |>
    summarise(pct = round(mean(easy) * 100)) |>
    pull(pct)
} else NA

cat(glue(
  "**{n_resp} test users** participated in a usability evaluation of the JPE Dashboard, ",
  "an interactive tool for exploring discrete water quality and fish genetics datasets. ",
  "Participants represented a range of professional roles, including staff from government ",
  "and academic agencies.\n\n",
  "The dashboard received positive usability ratings across key interface components:\n\n",
  "- Buttons and controls average score: **{avg_b}/5**\n",
  "- Visualizations average score: **{avg_v}/5**\n",
  "- Layout and organization average score: **{avg_l}/5**\n\n",
  "Initial navigation was rated positively, with {nav_summary}. ",
  "{if (!is.na(n_easy_tasks)) glue('Across four guided analytical tasks, {n_easy_tasks}% of responses indicated the task was completed easily.') else ''}\n\n",
  "Qualitative feedback highlighted four recurring themes: navigation and interface structure, ",
  "controls and filter interactions, data visualization and coverage, and system language and ",
  "messaging. Recommended improvements primarily relate to refining interactive controls, ",
  "clarifying how data is presented, and expanding functionality in the genetics module."
))
```

### Prioritized Recommendations

```{r recommendations}
recs <- tribble(
  ~Priority, ~Recommendation,
  "High",   "Recommendation #1 here",
  "High",   "Recommendation #2 here",
  "Medium", "Recommendation #3 here",
  "Medium", "Recommendation #4 here",
  "Low",    "Recommendation #5 here",
  "Low",    "Recommendation #6 here",
  "Low",    "Recommendation #7 here"
)

recs |>
  mutate(Priority = cell_spec(Priority, color = "white",
    background = case_when(
      Priority == "High"   ~ ROSE,
      Priority == "Medium" ~ AMBER,
      Priority == "Low"    ~ JADE), bold = TRUE)) |>
  kbl(escape = FALSE, col.names = c("Priority", "Recommendation"), align = "ll") |>
  kable_styling(bootstrap_options = c("hover", "condensed"), full_width = TRUE) |>
  column_spec(1, width = "7em", extra_css = "text-align:center") |>
  column_spec(2, width = "60em")
```

---

## Respondent Overview

### Roles


```{r role-chart}
#| results: asis

# define the lookup table
role_categories <- tribble(
  ~role_pattern,  ~role_category,
  "genetic specialist|genetic", "Environmental Scientist",
  "environmental scientist|scientist", "Environmental Scientist",
  "data manager|jpe data", "Data Manager",
  "consultant|design", "Dashboard / Data Analyst",
  "faculty|professor|academic|ucd", "Researcher / Academic"
)

# add role_category column to df
df <- df |>
  mutate(
    role_lower    = str_to_lower(role),
    role_category = map_chr(role_lower, function(x) {
      idx <- detect_index(role_categories$role_pattern,
                          ~ str_detect(x, regex(.x, ignore_case = TRUE)))
      if (idx > 0) role_categories$role_category[idx] else "Other"
    })
  )

# summarise and print
role_summary <- df |>
  count(role_category) |>
  arrange(desc(n))

cat("A total of **", nrow(df), "** individuals participated in this study. ",
    "Respondents represented a range of professional backgrounds:\n\n", sep = "")


# cat(glue("\n\nFive of {nrow(df)} respondents ",
#          "({round(mean(df$used_dashboard == 'Yes') * 100)}%) ",
#          "reported prior experience using other data dashboards.\n"))
```


### Dataset Familiarity

```{r familiarity-chart}
#| fig-cap: "Average self-reported familiarity (1 = not familiar, 3 = very familiar)"

fam_avail <- intersect(c("wq_familiar", "gen_familiar"), names(df))

if (length(fam_avail) > 0) {
  df |>
    select(any_of(c("wq_familiar", "gen_familiar"))) |>
    summarise(across(everything(), ~ mean(.x, na.rm = TRUE))) |>
    pivot_longer(everything(), names_to = "dataset", values_to = "avg") |>
    mutate(dataset = recode(dataset,
                            wq_familiar  = "Water Quality",
                            gen_familiar = "Fish Genetics")) |>
    ggplot(aes(avg, dataset, fill = dataset)) +
    geom_col(width = 0.4, show.legend = FALSE) +
    geom_text(aes(label = round(avg, 1)), hjust = -0.3,
              size = 4.5, fontface = "bold", color = INK) +
    scale_fill_manual(values = c("Water Quality" = OCEAN, "Fish Genetics" = JADE)) +
    scale_x_continuous(
      limits = c(0, 3.5),
      breaks = 1:3,
      labels = c("1\nNot familiar", "2\nSomewhat", "3\nVery familiar")
    ) +
    labs(title = "Average Dataset Familiarity", x = NULL, y = NULL)
}
```

### Prior Dashboard Experience

```{r dashboard-exp}
if ("used_dashboard" %in% names(df)) {
  df |>
    count(used_dashboard) |>
    mutate(`Percent of total participants` =
             scales::percent(n / sum(n), accuracy = 1)) |>
    select(used_dashboard, `Percent of total participants`) |>
    kbl(
      col.names = c("Used dashboards before?", "Percent of total participants"),
      align = "lr"
    ) |>
    kable_styling(
      bootstrap_options = c("striped", "hover", "condensed"),
      full_width = FALSE
    )
}
```

---

## Usability Ratings

### Intuitiveness

```{r intuit-bar}
#| fig-cap: "Average intuitiveness scores (1–5 scale)"
intuit_avail <- intersect(c("intuit_buttons", "intuit_viz", "intuit_layout"), names(df))

if (length(intuit_avail) > 0) {
  df |>
    summarise(across(any_of(intuit_avail), ~ mean(.x, na.rm = TRUE))) |>
    pivot_longer(everything(), names_to = "dim", values_to = "avg") |>
    mutate(
      dim   = recode(dim,
                intuit_buttons = "Buttons & Controls",
                intuit_viz     = "Visualizations",
                intuit_layout  = "Layout"),
      color = case_when(avg >= 4 ~ JADE, avg >= 3 ~ AMBER, .default = ROSE)
    ) |>
    ggplot(aes(fct_reorder(dim, avg), avg, fill = color)) +
    geom_col(width = 0.5, show.legend = FALSE) +
    geom_text(aes(label = round(avg, 1)), hjust = -0.3,
              size = 5, fontface = "bold", color = INK) +
    scale_fill_identity() +
    scale_y_continuous(limits = c(0, 5.8), breaks = 0:5) +
    coord_flip() +
    labs(title = "Average Intuitiveness Ratings",
         subtitle = "Green ≥ 4  |  Amber ≥ 3  |  Red < 3",
         x = NULL, y = "Average Score (1–5)")
}
```


---

## Task Completion

Participants completed four guided analytical tasks designed to assess the dashboard's 
ability to support typical data exploration workflows. These tasks required users to 
navigate the interface, filter and compare datasets, and interpret both water quality 
and fish genetics information across multiple sites and time periods. The table below 
describes each task, and the figure summarizes how participants rated the ease of 
completing them.

- **Task 1 (Water Quality):** Identify whether peaks in chlorophyll-a concentrations 
  occurred at San Pablo Bay sites (D41, D41A, NZ325) during 2022–2023.
- **Task 2 (Water Quality):** Determine whether changes in dissolved oxygen levels 
  correlate with shifts in water temperature across San Joaquin sites 
  (C10, D12, D16, D26, P8).
- **Task 3 (Fish Genetics):** Estimate the proportion of Chinook salmon samples on 
  Deer Creek classified as spring run in 2025.
- **Task 4 (Fish Genetics):** Identify which creek — Battle or Clear Creek — had a 
  greater number of early Greb 1L RoSA genotypes between 2023 and 2025.


```{r task-chart}
#| fig-cap: "Self-reported ease of completing four guided analytical tasks"
task_avail <- intersect(
  c("task_chla", "task_do_temp", "task_spring", "task_creek"), names(df))
task_labels <- c(
  task_chla    = "Task #1",
  task_do_temp = "Task #2",
  task_spring  = "Task #3",
  task_creek   = "Task #4"
)
if (length(task_avail) > 0) {
  df |>
    select(any_of(task_avail)) |>
    pivot_longer(everything(), names_to = "task", values_to = "response") |>
    filter(!is.na(response), str_length(response) > 1) |>
    mutate(
      difficulty = parse_task(response),
      task_label = recode(task, !!!task_labels[task_avail]),
      difficulty = factor(difficulty,
        levels = c("Easy", "Difficult", "Very Difficult", "Not answered"))
    ) |>
    filter(!is.na(difficulty)) |>
    count(task_label, difficulty) |>
    ggplot(aes(task_label, n, fill = difficulty)) +
    geom_col(width = 0.55) +
    scale_fill_manual(
      values = c(Easy = JADE, Difficult = AMBER,
                 `Very Difficult` = ROSE, `Not answered` = "#CBD5E1"),
      name = "Difficulty"
    ) +
    scale_y_continuous(breaks = scales::breaks_width(1),
                       expand = expansion(mult = c(0, 0.12))) +
    labs(title = "Task Completion Difficulty",
         subtitle = "Based on respondents' self-reported ease for use-case testing",
         x = NULL, y = "Respondents")
}
```

---

## Qualitative Feedback

Participants were asked to provide open-ended feedback following completion of the usability tasks. Responses addressed overall ease of use, navigation, clarity of visualizations, and potential areas for improvement. Individual comments were reviewed and organized into recurring themes, which are summarized below.

```{r qualitative-feedback, results='asis'}
# QUALITATIVE THEMES — reads classifications directly from Excel sheet

themes <- readxl::read_xlsx(here("feedback", "quantitative_themes.xlsx")) # I have added a v2 since I wanted to double-check that 

# Read the response + category block (rows 7–12, columns A–D)

coded_raw <- themes |> 
  slice(1:6) |>
  clean_names() |>
  rename(
    confusing    = 1,
    improvements = 2,
    add_comments = 3,
    categories   = 4
  ) |>
  mutate(respondent = row_number())   # anonymous ID — no role column

# Read curated quotes table (rows 16–20, columns A–B)
quotes_raw <- themes |> 
  row_to_names(row_number = 10) |>
  clean_names() |> 
  select(category, quote) |> 
  rename(theme = category, quotes_block = quote) 


# Reshape responses, one row per response × theme ───────────────

qual_cols <- c("confusing", "improvements", "add_comments")

col_labels <- c(
  confusing    = "Confusing / Difficult Elements",
  improvements = "Feature Requests & Improvements",
  add_comments = "Additional Comments"
)

# Long format: one row per respondent × question
qual_long <- coded_raw |>
  pivot_longer(
    cols      = all_of(qual_cols),
    names_to  = "question",
    values_to = "response"
  ) |>
  filter(!is.na(response), str_length(str_squish(response)) > 5) |>
  mutate(question_label = recode(question, !!!col_labels))

# Explode the comma-separated categories column into one row per theme
qual_exploded <- qual_long |>
  filter(!is.na(categories)) |>
  mutate(theme = str_split(categories, ",\\s*")) |>
  unnest(theme) |>
  mutate(theme = str_trim(theme))

# Consistent factor order matching your sheet
theme_levels <- c(
  "Navigation/Interface",
  "Controls/Filters",
  "Visualization/Data",
  "Language/Messages")

theme_colors <- c(
  "Navigation/Interface" = "#2563EB",
  "Controls/Filters"     = "#F43F5E",
  "Visualization/Data"   = "#10B981",
  "Language/Messages"    = "#F59E0B")

qual_exploded <- qual_exploded |>
  mutate(theme = factor(theme, levels = theme_levels))


# Parse curated quotes into one row per quote ───────────────

quotes_long <- quotes_raw |>
  mutate(
    quote = str_split(quotes_block, "\n\n")
  ) |>
  unnest(quote) |>
  mutate(
    quote = str_trim(quote),
    # Remove outer wrapping quotes (curly and straight)
    quote = str_remove_all(quote, '^[\u201c\u201d"]+|[\u201c\u201d"]+$'),
    # Remove any parenthetical notes like "(Partly layout / partly feedback)"
    quote = str_remove(quote, "^\\(.*?\\)\\s*"),
    # Clean up any leftover leading/trailing punctuation
    quote = str_trim(quote),
    theme = factor(theme, levels = theme_levels)
  ) |>
  filter(str_length(quote) > 5)


# One styled callout card per theme ───────

for (grp in theme_levels) {

  quotes_for_theme <- quotes_long |>
    dplyr::filter(theme == grp) |>
    dplyr::pull(quote) |>
    stringr::str_split("\\r\\n\\r\\n|\\n\\n") |>
    unlist() |>
    stringr::str_squish() |>
    stringr::str_remove_all('^“|”$|^"|"$')

  quotes_for_theme <- quotes_for_theme[stringr::str_length(quotes_for_theme) > 0]

  if (length(quotes_for_theme) == 0) next

  color <- theme_colors[[grp]]

  bullet_quotes <- paste0(
    "<li>&ldquo;", quotes_for_theme, "&rdquo;</li>",
    collapse = "\n"
  )

  cat(glue::glue('
<div style="border-left: 4px solid {color}; padding: 10px 16px; margin-bottom: 20px; background: #FAFAFA;">
  <strong>{grp}</strong>
  <ul style="margin-top: 8px; margin-bottom: 0;">
    {bullet_quotes}
  </ul>
</div>

'))
}
```



---


## Appendix

### Appendix A: Action Item Checklist 

This checklist summarizes recommended improvements identified from participant feedback, 
organized by theme. Items can be used to prioritize development tasks.

#### Navigation / Interface

- [ ] Relocate the Water Quality and Genetics tabs from the top of the page to a more
      conventional location (e.g. sidebar or persistent top nav) - *LOW*
- [ ] Restructure the Overview page — replace long paragraphs with concise topic 
      sentences and bullet points - *LOW*
- [ ] Add a clear section heading that states the dashboard purpose upfront - *LOW*
- [ ] Sort station names alphabetically with station ID in parentheses - *LOW*
- [ ] Remove auto-zoom to selected sites  - *LOW*
- [ ] Add color differentiation for selected sites that are active vs inactive  - *LOW*


#### Controls / Filters

- [ ] Replace the year slider with checkboxes or a date range picker - *HIGH*
- [ ] Remove tick marks between years on the year toggle to prevent user confusion - *MEDIUM*
- [ ] Relabel the year toggle to say "Water Year" instead of "Year" - *LOW*
- [ ] Fix the site/analyte dropdowns so options do not disappear unexpectedly
      when switching between sites - *LOW*
- [ ] Allow users to select within-year date ranges for more granular filtering - *LOW*

#### Visualization / Data

- [ ] Add river labels to the water quality map - *MEDIUM*
- [ ] Add city labels to the water quality map - *MEDIUM*
- [ ] Add water quality monitoring stations upstream of Sacramento at Hood - *?*
- [ ] Display raw fish counts alongside proportions in the genetics module - *LOW (ADD EXPLANATION OF TOOLTIP)*
- [ ] Add daily or event-level catch data rather than annual totals only - *?*
- [ ] Add water year type and flow/export data overlays to the genetics dashboard - *MEDUIM*
- [ ] Build a comparison view for field length-at-date assignments vs. genetic
      run assignments - *MEDIUM*
- [ ] Replace the single large box plot with time-series summary plots
      (monthly or annual) - *?*
- [ ] Add an option to toggle between average, maximum, minimum, and
      7-day average values - *NA (EXPLAIN WATER DATA COLLECTED ONCE A MONTH)*
- [ ] Add links or connections to EDI pages for metadata, citations, and downloads - *LOW*
- [ ] Clarify the difference between surface and bottom sensor measurements
      (e.g. "Dissolved Oxygen" vs. "(bottom) Dissolved Oxygen") - *MEDIUM*
- [ ] Indicate whether more historical data or future data will be available - *LOW*

#### Language / Messages

- [ ] Clarify what "inactive" means in the site selector (currently inactive vs.
      inactive across full time range) - *LOW*
- [ ] Add a chart subtitle or title note specifying what data is shown
      (e.g. "Average daily concentrations") - *MEDUIM*
- [ ] Add a label reminding users that water quality data shown is
      discrete monthly samples - *LOW - ADD EXPLANATION*

### User Survey

[Link to survey]

```{r raw-data}
# df |>
#   select(-matches("email|timestamp", ignore.case = TRUE)) |>
#   head(10) |>
#   kbl() |>
#   kable_styling(
#     bootstrap_options = c("striped", "hover", "condensed", "responsive"),
#     full_width = TRUE, font_size = 11
#   ) |>
#   scroll_box(width = "100%", height = "360px")
```

