---
title: "Downstream Dashboard Dashboard — User Research Feedback Report"
subtitle: "Compiled from [Google Forms Survey Responses](https://docs.google.com/forms/d/e/1FAIpQLSe3M5DvdxeucI8y7Zg3buI1ZdN5KMstl5CULJObjRpkzUfg8Q/viewform?usp=header). \n\n **DRAFT - in development**"
date: today
date-format: "MMMM YYYY"

format:
  html:
    theme: cosmo
    checkbox: true
    toc: true
    toc-depth: 3
    toc-title: "Contents"
    toc-location: left
    number-sections: true
    fig-width: 8
    fig-height: 5
    fig-dpi: 150
    embed-resources: true        # single self-contained HTML file
    code-fold: true              # hide code by default, togglable
    code-summary: "Show code"

  pdf:
    toc: true
    number-sections: true
    colorlinks: true
    fig-pos: "H"

execute:
  echo: false
  warning: false
  message: false
  cache: false   # set true to avoid re-fetching sheet on every render
---

```{r setup}
#| label: setup
#| include: false

# ── Install / load packages ──────────────────────────────────────────────────
library("googlesheets4")
library("tidyverse")
library("scales")
library("knitr")
library("kableExtra")
library("fmsb")
library("glue")
library("janitor")
library("lubridate")

# ── Design tokens ────────────────────────────────────────────────────────────
OCEAN <- "#2563EB"; TEAL <- "#0EA5E9"; JADE <- "#10B981"
AMBER <- "#F59E0B"; ROSE <- "#F43F5E"; SLATE <- "#64748B"; INK <- "#1A1F2E"

# ── ggplot theme ─────────────────────────────────────────────────────────────
theme_survey <- function(base_size = 12) {
  theme_minimal(base_size = base_size) %+replace% theme(
    plot.title       = element_text(face = "bold", size = base_size + 2,
                                    color = INK, hjust = 0),
    plot.subtitle    = element_text(size = base_size - 1, color = SLATE,
                                    hjust = 0, margin = margin(b = 8)),
    axis.text        = element_text(color = SLATE, size = base_size - 2),
    axis.title       = element_text(color = SLATE, size = base_size - 1),
    panel.grid.major = element_line(color = "#E2E8F0"),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background  = element_rect(fill = "white", color = NA),
    legend.position  = "bottom",
    legend.title     = element_text(face = "bold", size = base_size - 1),
    strip.text       = element_text(face = "bold", color = INK)
  )
}
theme_set(theme_survey())
```

```{r load-data}
#| label: load-data
#| include: false

# ╔══════════════════════════════════════════════════════════════════════════╗
# ║  CONFIGURE YOUR GOOGLE SHEETS CONNECTION                                ║
# ╠══════════════════════════════════════════════════════════════════════════╣
# ║  Step 1 — Paste your Sheet URL or Sheet ID below.                       ║
# ║    • The URL looks like:                                                 ║
# ║      https://docs.google.com/spreadsheets/d/SHEET_ID/edit               ║
# ║    • Or just use the SHEET_ID string directly.                          ║
# ║                                                                          ║
# ║  Step 2 — Choose AUTH_MODE:                                              ║
# ║    "public"      Sheet is shared "Anyone with link can view"            ║
# ║                  → no login needed, works in CI/CD pipelines             ║
# ║    "interactive" Opens a browser popup to sign in once;                 ║
# ║                  token is cached at ~/.cache/gargle/                     ║
# ║    "service"     Use a service-account JSON key (headless/server use)   ║
# ║                  → set SERVICE_ACCOUNT_JSON path below                  ║
# ╚══════════════════════════════════════════════════════════════════════════╝

SHEET_URL            <- "https://docs.google.com/spreadsheets/d/1wKb_0v9vsbBZdtfBAZeUZTlZqvGR9YnBmRjPCE5Knq8/edit?gid=1013775912#gid=1013775912"
AUTH_MODE            <- "public"      # "public" | "interactive" | "service"
SERVICE_ACCOUNT_JSON <- NULL          # e.g. "credentials/sa_key.json"
SHEET_TAB            <- 1             # 1 = first tab, or use tab name as string

# ── Authenticate ─────────────────────────────────────────────────────────────
if (AUTH_MODE == "public") {
  gs4_deauth()
} else if (AUTH_MODE == "interactive") {
  gs4_auth()
} else if (AUTH_MODE == "service") {
  gs4_auth(path = SERVICE_ACCOUNT_JSON)
}

# ── Read sheet ────────────────────────────────────────────────────────────────
raw <- read_sheet(SHEET_URL, sheet = SHEET_TAB)

# ── Clean column names ────────────────────────────────────────────────────────
df <- raw |>
  clean_names() |>
  mutate(across(where(is.character), str_squish))

# ── Fuzzy-match key columns → standard names ──────────────────────────────────
# Adjust the regex patterns here if your form question wording differs.
safe_rename <- function(df, new_name, pattern) {
  matched <- names(df)[str_detect(names(df), regex(pattern, ignore_case = TRUE))]
  if (length(matched) > 0 && !new_name %in% names(df))
    df <- df |> rename(!!new_name := all_of(matched[[1]]))
  df
}

df <- df |>
  safe_rename("role",           "role|position|title") |>
  safe_rename("wq_familiar",    "water.quality.*familiar|familiar.*water") |>
  safe_rename("gen_familiar",   "genetic.*familiar|familiar.*genetic") |>
  safe_rename("used_dashboard", "used.*dashboard|dashboard.*before") |>
  safe_rename("nav_ease",       "easy.*navig|navig.*easy|first.use") |>
  safe_rename("intuit_buttons", "buttons.*controls|controls.*buttons") |>
  safe_rename("intuit_viz",     "visualization") |>
  safe_rename("intuit_layout",  "layout") |>
  safe_rename("confusing",      "confus|difficult.*use") |>
  safe_rename("improvements",   "improve|features") |>
  safe_rename("add_comments",   "additional.*comment|^comment") |>
  safe_rename("task_chla",      "chlorophyll|chla") |>
  safe_rename("task_do_temp",   "dissolved.*oxygen|do.*temp") |>
  safe_rename("task_spring",    "spring.*run|chinook") |>
  safe_rename("task_creek",     "battle.*creek|creek.*battle|genotype")

# ── Coerce ratings to numeric ─────────────────────────────────────────────────
rating_cols <- c("wq_familiar", "gen_familiar", "nav_ease",
                 "intuit_buttons", "intuit_viz", "intuit_layout")

df <- df |>
  mutate(across(any_of(rating_cols), as.numeric))

n_resp <- nrow(df)
```

```{r helpers}
#| include: false

# Classify task response text as Easy / Difficult / Very Difficult
parse_task <- function(x) {
  x_low <- str_to_lower(x)
  case_when(
    str_detect(x_low, "very.difficult|very hard|couldn|not able") ~ "Very Difficult",
    str_detect(x_low, "difficult|hard")                           ~ "Difficult",
    str_detect(x_low, "easy|straightforward|simple|no problem")   ~ "Easy",
    is.na(x) | str_length(x) < 2                                  ~ NA_character_,
    .default = "Not answered"
  )
}
```

---

## Executive Summary

```{r exec-summary}
#| results: asis

avg_b <- if ("intuit_buttons" %in% names(df)) round(mean(df$intuit_buttons, na.rm = TRUE), 1) else "N/A"
avg_v <- if ("intuit_viz"     %in% names(df)) round(mean(df$intuit_viz,     na.rm = TRUE), 1) else "N/A"
avg_l <- if ("intuit_layout"  %in% names(df)) round(mean(df$intuit_layout,  na.rm = TRUE), 1) else "N/A"
n_resp <- nrow(df)

# Precompute nav ease summary
nav_summary <- if ("nav_ease" %in% names(df)) {
  n_easy <- sum(df$nav_ease <= 2, na.rm = TRUE)
  glue("{n_easy} of {n_resp} participants rated it easy or very easy")
} else "straightforward"

# Precompute task completion summary
task_cols_avail <- intersect(c("task_chla", "task_do_temp", "task_spring", "task_creek"), names(df))

n_easy_tasks <- if (length(task_cols_avail) > 0) {
  df |>
    select(all_of(task_cols_avail)) |>
    pivot_longer(everything(), values_to = "response") |>
    filter(!is.na(response), str_length(response) > 1) |>
    mutate(easy = str_detect(str_to_lower(response), "easy|straightforward|simple")) |>
    summarise(pct = round(mean(easy) * 100)) |>
    pull(pct)
} else NA

cat(glue(
  "**{n_resp} test users** participated in a usability evaluation of the JPE Dashboard, ",
  "an interactive tool for exploring discrete water quality and fish genetics datasets. ",
  "Participants represented a range of professional roles, including staff from government ",
  "and academic agencies.\n\n",
  "The dashboard received positive usability ratings across key interface components:\n\n",
  "- Buttons and controls average score: **{avg_b}/5**\n",
  "- Visualizations average score: **{avg_v}/5**\n",
  "- Layout and organization average score: **{avg_l}/5**\n\n",
  "Initial navigation was rated positively, with {nav_summary}. ",
  "{if (!is.na(n_easy_tasks)) glue('Across four guided analytical tasks, {n_easy_tasks}% of responses indicated the task was completed easily.') else ''}\n\n",
  "Qualitative feedback highlighted four recurring themes: navigation and interface structure, ",
  "controls and filter interactions, data visualization and coverage, and system language and ",
  "messaging. Recommended improvements primarily relate to refining interactive controls, ",
  "clarifying how data is presented, and expanding functionality in the genetics module."
))
```

### Prioritized Recommendations

```{r recommendations}
recs <- tribble(
  ~Priority, ~Recommendation,
  "High",   "Recommendation #1 here",
  "High",   "Recommendation #2 here",
  "Medium", "Recommendation #3 here",
  "Medium", "Recommendation #4 here",
  "Low",    "Recommendation #5 here",
  "Low",    "Recommendation #6 here",
  "Low",    "Recommendation #7 here"
)

recs |>
  mutate(Priority = cell_spec(Priority, color = "white",
    background = case_when(
      Priority == "High"   ~ ROSE,
      Priority == "Medium" ~ AMBER,
      Priority == "Low"    ~ JADE), bold = TRUE)) |>
  kbl(escape = FALSE, col.names = c("Priority", "Recommendation"), align = "ll") |>
  kable_styling(bootstrap_options = c("hover", "condensed"), full_width = TRUE) |>
  column_spec(1, width = "7em", extra_css = "text-align:center") |>
  column_spec(2, width = "60em")
```

---

## Respondent Overview

### Roles


```{r role-chart}
#| results: asis

# Step 1 — define the lookup table
role_categories <- tribble(
  ~role_pattern,  ~role_category,
  "genetic specialist|genetic", "Environmental Scientist",
  "environmental scientist|scientist", "Environmental Scientist",
  "data manager|jpe data", "Data Manager",
  "consultant|design", "Dashboard / Data Analyst",
  "faculty|professor|academic|ucd", "Researcher / Academic"
)

# Step 2 — add role_category column to df
df <- df |>
  mutate(
    role_lower    = str_to_lower(role),
    role_category = map_chr(role_lower, function(x) {
      idx <- detect_index(role_categories$role_pattern,
                          ~ str_detect(x, regex(.x, ignore_case = TRUE)))
      if (idx > 0) role_categories$role_category[idx] else "Other"
    })
  )

# Step 3 — summarise and print
role_summary <- df |>
  count(role_category) |>
  arrange(desc(n))

cat("A total of **", nrow(df), "** individuals participated in this study. ",
    "Respondents represented a range of professional backgrounds:\n\n", sep = "")

# role_summary |>
#   pwalk(function(role_category, n) {
#     pct <- round((n / nrow(df)) * 100)
#     cat(glue("- **{role_category}** ({n} respondent{if (n > 1) 's' else ''}; {pct}% of total)\n\n"))
#   })

cat(glue("\n\nFive of {nrow(df)} respondents ",
         "({round(mean(df$used_dashboard == 'Yes') * 100)}%) ",
         "reported prior experience using other data dashboards.\n"))
```


### Dataset Familiarity

```{r familiarity-chart}
#| fig-cap: "Average self-reported familiarity (1 = not familiar, 3 = very familiar)"

fam_avail <- intersect(c("wq_familiar", "gen_familiar"), names(df))

if (length(fam_avail) > 0) {
  df |>
    select(any_of(c("wq_familiar", "gen_familiar"))) |>
    summarise(across(everything(), ~ mean(.x, na.rm = TRUE))) |>
    pivot_longer(everything(), names_to = "dataset", values_to = "avg") |>
    mutate(dataset = recode(dataset,
                            wq_familiar  = "Water Quality",
                            gen_familiar = "Fish Genetics")) |>
    ggplot(aes(avg, dataset, fill = dataset)) +
    geom_col(width = 0.4, show.legend = FALSE) +
    geom_text(aes(label = round(avg, 1)), hjust = -0.3,
              size = 4.5, fontface = "bold", color = INK) +
    scale_fill_manual(values = c("Water Quality" = OCEAN, "Fish Genetics" = JADE)) +
    scale_x_continuous(
      limits = c(0, 3.5),
      breaks = 1:3,
      labels = c("1\nNot familiar", "2\nSomewhat", "3\nVery familiar")
    ) +
    labs(title = "Average Dataset Familiarity", x = NULL, y = NULL)
}
```

### Prior Dashboard Experience

```{r dashboard-exp}
if ("used_dashboard" %in% names(df)) {
  df |>
    count(used_dashboard) |>
    mutate(`Percent of total participants` =
             scales::percent(n / sum(n), accuracy = 1)) |>
    select(used_dashboard, `Percent of total participants`) |>
    kbl(
      col.names = c("Used dashboards before?", "Percent of total participants"),
      align = "lr"
    ) |>
    kable_styling(
      bootstrap_options = c("striped", "hover", "condensed"),
      full_width = FALSE
    )
}
```

---

## Usability Ratings


```{r nav-ease}
### First-Use Navigation Ease
# #| fig-cap: "Distribution of first-use navigation ease scores"
# if ("nav_ease" %in% names(df)) {
#   df |>
#     filter(!is.na(nav_ease)) |>
#     count(nav_ease) |>
#     mutate(label = case_when(
#       nav_ease == 1 ~ "1 – Very Easy",
#       nav_ease == 2 ~ "2 – Easy",
#       nav_ease == 3 ~ "3 – Moderate",
#       nav_ease == 4 ~ "4 – Difficult",
#       nav_ease == 5 ~ "5 – Very Difficult",
#       .default = as.character(nav_ease)
#     )) |>
#     ggplot(aes(fct_reorder(label, nav_ease), n)) +
#     geom_col(fill = TEAL, width = 0.5) +
#     geom_text(aes(label = n), vjust = -0.4, size = 4.5,
#               fontface = "bold", color = INK) +
#     scale_y_continuous(expand = expansion(mult = c(0, 0.2)),
#                        breaks = scales::breaks_width(1)) +
#     labs(title = "Ease of First-Time Navigation",
#          subtitle = "All respondents rated navigation easy or very easy",
#          x = NULL, y = "Respondents")
# }
```

### Intuitiveness

```{r intuit-bar}
#| fig-cap: "Average intuitiveness scores (1–5 scale)"
intuit_avail <- intersect(c("intuit_buttons", "intuit_viz", "intuit_layout"), names(df))

if (length(intuit_avail) > 0) {
  df |>
    summarise(across(any_of(intuit_avail), ~ mean(.x, na.rm = TRUE))) |>
    pivot_longer(everything(), names_to = "dim", values_to = "avg") |>
    mutate(
      dim   = recode(dim,
                intuit_buttons = "Buttons & Controls",
                intuit_viz     = "Visualizations",
                intuit_layout  = "Layout"),
      color = case_when(avg >= 4 ~ JADE, avg >= 3 ~ AMBER, .default = ROSE)
    ) |>
    ggplot(aes(fct_reorder(dim, avg), avg, fill = color)) +
    geom_col(width = 0.5, show.legend = FALSE) +
    geom_text(aes(label = round(avg, 1)), hjust = -0.3,
              size = 5, fontface = "bold", color = INK) +
    scale_fill_identity() +
    scale_y_continuous(limits = c(0, 5.8), breaks = 0:5) +
    coord_flip() +
    labs(title = "Average Intuitiveness Ratings",
         subtitle = "Green ≥ 4  |  Amber ≥ 3  |  Red < 3",
         x = NULL, y = "Average Score (1–5)")
}
```



```{r rating-table}
# ### Ratings Summary Table
# if (length(intuit_avail) > 0) {
#   df |>
#     select(any_of(intuit_avail)) |>
#     summarise(across(everything(), list(
#       Mean = ~ round(mean(.x, na.rm = TRUE), 2),
#       SD   = ~ round(sd(.x,   na.rm = TRUE), 2),
#       Min  = ~ min(.x, na.rm = TRUE),
#       Max  = ~ max(.x, na.rm = TRUE)
#     ))) |>
#     pivot_longer(everything(),
#                  names_to  = c("Dimension", ".value"),
#                  names_sep = "_") |>
#     mutate(Dimension = recode(Dimension,
#       intuit.buttons = "Buttons & Controls",
#       intuit.viz     = "Visualizations",
#       intuit.layout  = "Layout"
#     )) |>
#     kbl(align = "lrrrr") |>
#     kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
#                   full_width = FALSE) |>
#     column_spec(2, bold = TRUE, color = OCEAN)
# }
```

---

## Task Completion

Participants completed four guided analytical tasks designed to assess the dashboard’s ability to support typical data exploration workflows. These tasks required users to navigate the interface, filter and compare datasets, and interpret both water quality, and fish genetics information across multiple sites and time periods

```{r task-chart}
#| fig-cap: "Self-reported ease of completing four guided analytical tasks"
task_avail <- intersect(
  c("task_chla", "task_do_temp", "task_spring", "task_creek"), names(df))

task_labels <- c(
  task_chla    = "Task #1",
  task_do_temp = "Task #2",
  task_spring  = "Task #3",
  task_creek   = "Task #4"
)

if (length(task_avail) > 0) {
  df |>
    select(any_of(task_avail)) |>
    pivot_longer(everything(), names_to = "task", values_to = "response") |>
    filter(!is.na(response), str_length(response) > 1) |>
    mutate(
      difficulty = parse_task(response),
      task_label = recode(task, !!!task_labels[task_avail]),
      difficulty = factor(difficulty,
        levels = c("Easy", "Difficult", "Very Difficult", "Not answered"))
    ) |>
    filter(!is.na(difficulty)) |>
    count(task_label, difficulty) |>
    ggplot(aes(task_label, n, fill = difficulty)) +
    geom_col(width = 0.55) +
    scale_fill_manual(
      values = c(Easy = JADE, Difficult = AMBER,
                 `Very Difficult` = ROSE, `Not answered` = "#CBD5E1"),
      name = "Difficulty"
    ) +
    scale_y_continuous(breaks = scales::breaks_width(1),
                       expand = expansion(mult = c(0, 0.12))) +
    labs(title = "Task Completion Difficulty",
         subtitle = "Based on respondents' self-reported ease for use-case testing",
         x = NULL, y = "Respondents")
}
```

---

## Qualitative Feedback

Participants were asked to provide open-ended feedback following completion of the usability tasks. Responses addressed overall ease of use, navigation, clarity of visualizations, and potential areas for improvement. Individual comments were reviewed and organized into recurring themes, which are summarized below.

```{r qualitative-feedback, results='asis'}
# =============================================================================
# QUALITATIVE THEMES — reads classifications directly from your Excel sheet
# No keyword matching. Uses exactly what you coded in Quantitative_themes tab.
# =============================================================================

# ── SECTION A: Load your pre-coded classifications from Excel ─────────────────

library(readxl)
library(here)

themes <- readxl::read_xlsx(here("feedback", "quantitative_themes.xlsx"))

# Read the response + category block (rows 7–12, columns A–D)
# Adjust the path to wherever your Excel file lives
coded_raw <- themes |> 
  slice(1:6) |>
  clean_names() |>
  rename(
    confusing    = 1,
    improvements = 2,
    add_comments = 3,
    categories   = 4
  ) |>
  mutate(respondent = row_number())   # anonymous ID — no role column

# Read your curated quotes table (rows 16–20, columns A–B)
quotes_raw <- themes |> 
  row_to_names(row_number = 10) |>
  clean_names() |> 
  select(category, quote) |> 
  rename(theme = category, quotes_block = quote) 


# ── SECTION B: Reshape responses — one row per response × theme ───────────────

qual_cols <- c("confusing", "improvements", "add_comments")

col_labels <- c(
  confusing    = "Confusing / Difficult Elements",
  improvements = "Feature Requests & Improvements",
  add_comments = "Additional Comments"
)

# Long format: one row per respondent × question
qual_long <- coded_raw |>
  pivot_longer(
    cols      = all_of(qual_cols),
    names_to  = "question",
    values_to = "response"
  ) |>
  filter(!is.na(response), str_length(str_squish(response)) > 5) |>
  mutate(question_label = recode(question, !!!col_labels))

# Explode the comma-separated categories column into one row per theme
qual_exploded <- qual_long |>
  filter(!is.na(categories)) |>
  mutate(theme = str_split(categories, ",\\s*")) |>
  unnest(theme) |>
  mutate(theme = str_trim(theme))

# Consistent factor order matching your sheet
theme_levels <- c(
  "Navigation/Interface",
  "Controls/Filters",
  "Visualization/Data",
  "Language/Messages"
)

theme_colors <- c(
  "Navigation/Interface" = "#2563EB",
  "Controls/Filters"     = "#F43F5E",
  "Visualization/Data"   = "#10B981",
  "Language/Messages"    = "#F59E0B"
)

qual_exploded <- qual_exploded |>
  mutate(theme = factor(theme, levels = theme_levels))


# ── SECTION C: Parse your curated quotes into one row per quote ───────────────

quotes_long <- quotes_raw |>
  mutate(
    quote = str_split(quotes_block, "\n\n")
  ) |>
  unnest(quote) |>
  mutate(
    quote = str_trim(quote),
    # Remove outer wrapping quotes (curly and straight)
    quote = str_remove_all(quote, '^[\u201c\u201d"]+|[\u201c\u201d"]+$'),
    # Remove any parenthetical notes like "(Partly layout / partly feedback)"
    quote = str_remove(quote, "^\\(.*?\\)\\s*"),
    # Clean up any leftover leading/trailing punctuation
    quote = str_trim(quote),
    theme = factor(theme, levels = theme_levels)
  ) |>
  filter(str_length(quote) > 5)

# ── SECTION D: Theme frequency bar chart (the one you liked) ─────────────────

# qual_exploded |>
#   count(theme) |>
#   mutate(theme = fct_reorder(theme, n)) |>
#   ggplot(aes(n, theme, fill = as.character(theme))) +
#   geom_col(width = 0.6, show.legend = FALSE) +
#   geom_text(aes(label = n), hjust = -0.3,
#             size = 4.5, fontface = "bold", color = INK) +
#   scale_fill_manual(values = theme_colors) +
#   scale_x_continuous(
#     expand = expansion(mult = c(0, 0.2)),
#     breaks = scales::breaks_width(1)
#   ) +
#   labs(
#     title    = "Qualitative Feedback by Theme",
#     subtitle = "Number of responses coded to each theme (responses may appear in multiple themes)",
#     x        = "Number of coded responses",
#     y        = NULL
#   )


# ── SECTION E: Quoted evidence panel — one block per theme ───────────────────
# Displays your curated quotes under each theme heading, styled as a table.
# No respondent info shown — fully anonymous.

# quotes_long |>
#   arrange(theme) |>
#   group_by(theme) |>
#   mutate(quote = paste0("\u201c", quote, "\u201d")) |>   # add typographic quotes
#   summarise(quotes_combined = paste(quote, collapse = "\n\n"), .groups = "drop") |>
#   kbl(
#     col.names = c("Theme", "Representative Quotes"),
#     align     = "ll"
#   ) |>
#   kable_styling(
#     bootstrap_options = c("hover", "condensed"),
#     full_width        = TRUE
#   ) |>
#   column_spec(1, bold = TRUE, width = "16em",
#               color = "white",
#               background = theme_colors[as.character(quotes_long$theme |> unique() |> sort())]) |>
#   column_spec(2, width = "64em",
#               extra_css = "white-space: pre-wrap; line-height: 1.8;")


# ── SECTION F: One styled callout card per theme (alternative to table) ───────
# This version renders each theme as its own mini-section with a colored
# header and indented quotes — better for HTML output.
# Wrap this chunk in results: asis in your Quarto doc.

for (grp in theme_levels) {

  quotes_for_theme <- quotes_long |>
    dplyr::filter(theme == grp) |>
    dplyr::pull(quote) |>
    stringr::str_split("\\r\\n\\r\\n|\\n\\n") |>
    unlist() |>
    stringr::str_squish() |>
    stringr::str_remove_all('^“|”$|^"|"$')

  quotes_for_theme <- quotes_for_theme[stringr::str_length(quotes_for_theme) > 0]

  if (length(quotes_for_theme) == 0) next

  color <- theme_colors[[grp]]

  bullet_quotes <- paste0(
    "<li>&ldquo;", quotes_for_theme, "&rdquo;</li>",
    collapse = "\n"
  )

  cat(glue::glue('
<div style="border-left: 4px solid {color}; padding: 10px 16px; margin-bottom: 20px; background: #FAFAFA;">
  <strong>{grp}</strong>
  <ul style="margin-top: 8px; margin-bottom: 0;">
    {bullet_quotes}
  </ul>
</div>

'))
}
```













```{r confusing-table}

### Confusing or Difficult Elements

# pending: summary of feedback

# if ("confusing" %in% names(df)) {
#   df |>
#     select(any_of("role"), confusing) |>
#     filter(!is.na(confusing), str_length(confusing) > 3) |>
#     mutate(confusing = str_trunc(confusing, 450)) |>
#     kbl(col.names = c("Role", "Response"), align = "ll") |>
#     kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE) |>
#     column_spec(1, width = "18em", italic = TRUE, color = SLATE) |>
#     column_spec(2, width = "60em")
# }
```


```{r improvements-table}

### Feature Requests & Improvements

# pending: summary of feedback

# if ("improvements" %in% names(df)) {
#   df |>
#     select(any_of("role"), improvements) |>
#     filter(!is.na(improvements), str_length(improvements) > 3) |>
#     mutate(improvements = str_trunc(improvements, 450)) |>
#     kbl(col.names = c("Role", "Response"), align = "ll") |>
#     kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE) |>
#     column_spec(1, width = "18em", italic = TRUE, color = SLATE) |>
#     column_spec(2, width = "60em")
# }
```


```{r comments-table}
### Additional Comments
# if ("add_comments" %in% names(df)) {
#   df |>
#     select(any_of("role"), add_comments) |>
#     filter(!is.na(add_comments), str_length(add_comments) > 3) |>
#     mutate(add_comments = str_trunc(add_comments, 450)) |>
#     kbl(col.names = c("Role", "Response"), align = "ll") |>
#     kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE) |>
#     column_spec(1, width = "18em", italic = TRUE, color = SLATE) |>
#     column_spec(2, width = "60em")
# }
```



```{r themes-table}
### Recurring Themes
# themes_df <- tribble(
#   ~Theme,                           ~Frequency, ~Sentiment, ~Summary,
#   "Navigation & Controls Friction",  "4 / 6",  "Mixed",    "Tab placement, year slider, and disappearing dropdowns caused friction.",
#   "Data Context & Transparency",     "3 / 6",  "Neutral",  "Requests for aggregation labels, persistent tooltips, and inactive-site messages.",
#   "Genetics Module Expansion",       "3 / 6",  "Positive", "Desire for flow overlays, daily-level counts, raw fish numbers, and upstream stations.",
#   "Verbose Overview Page",           "2 / 6",  "Neutral",  "Overview text is too dense; users prefer concise headings and bullet structure.",
#   "Strong Overall Satisfaction",     "4 / 6",  "Positive", "Multiple respondents praised multi-site comparison and overall ease of use."
# )
# 
# themes_df |>
#   mutate(Sentiment = cell_spec(Sentiment, color = "white",
#     background = case_when(
#       Sentiment == "Positive" ~ JADE,
#       Sentiment == "Negative" ~ ROSE,
#       .default = SLATE
#     ), bold = TRUE)) |>
#   kbl(escape = FALSE, align = "llll") |>
#   kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE) |>
#   column_spec(1, bold = TRUE, width = "22em") |>
#   column_spec(2, width = "8em", extra_css = "text-align:center") |>
#   column_spec(3, width = "9em", extra_css = "text-align:center")
```

---


## Appendix

### Appendix A: Action Item Checklist 

This checklist summarizes recommended improvements identified from participant feedback, 
organized by theme. Items can be used to prioritize development tasks.

#### Navigation / Interface

- [ ] Relocate the Water Quality and Genetics tabs from the top of the page to a more
      conventional location (e.g. sidebar or persistent top nav)
- [ ] Restructure the Overview page — replace long paragraphs with concise topic
      sentences and bullet points
- [ ] Add a clear section heading that states the dashboard purpose upfront
- [ ] Sort station names alphabetically with station ID in parentheses

#### Controls / Filters

- [ ] Replace the year slider with checkboxes or a date range picker
- [ ] Remove tick marks between years on the year toggle to prevent user confusion
- [ ] Relabel the year toggle to say "Water Year" instead of "Year"
- [ ] Fix the site/analyte dropdowns so options do not disappear unexpectedly
      when switching between sites
- [ ] Allow users to select within-year date ranges for more granular filtering

#### Visualization / Data

- [ ] Add river labels to the water quality map
- [ ] Add water quality monitoring stations upstream of Sacramento at Hood
- [ ] Display raw fish counts alongside proportions in the genetics module
- [ ] Add daily or event-level catch data rather than annual totals only
- [ ] Add water year type and flow/export data overlays to the genetics dashboard
- [ ] Build a comparison view for field length-at-date assignments vs. genetic
      run assignments
- [ ] Replace the single large box plot with time-series summary plots
      (monthly or annual)
- [ ] Add an option to toggle between average, maximum, minimum, and
      7-day average values
- [ ] Add links or connections to EDI pages for metadata, citations, and downloads
- [ ] Clarify the difference between surface and bottom sensor measurements
      (e.g. "Dissolved Oxygen" vs. "(bottom) Dissolved Oxygen")
- [ ] Indicate whether more historical data or future data will be available

#### Language / Messages

- [ ] Add a "No data available" message when a user selects an inactive monitoring site
- [ ] Clarify what "inactive" means in the site selector (currently inactive vs.
      inactive across full time range)
- [ ] Add a persistent data exclusion note at the bottom of the plot when
      multiple locations are selected
- [ ] Add a chart subtitle or title note specifying what data is shown
      (e.g. "Average daily concentrations")
- [ ] Add a label reminding users that water quality data shown is
      discrete monthly samples

### User Survey

[Link to survey]

```{r raw-data}
# df |>
#   select(-matches("email|timestamp", ignore.case = TRUE)) |>
#   head(10) |>
#   kbl() |>
#   kable_styling(
#     bootstrap_options = c("striped", "hover", "condensed", "responsive"),
#     full_width = TRUE, font_size = 11
#   ) |>
#   scroll_box(width = "100%", height = "360px")
```

